---
title: "My Tools Collection"
author: "Yijun Xie"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{My Tools Collection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### This is a collection of my commonly used functions.

This package includes 4 functions:

## chi.bar()
This is a function to calculate the coefficient of tail independence. It will return the estimated value as well as the confidence interval of the estimated coefficient of tail independence.
An example can be find here:
```{r}
chi.bar(rnorm(1000),q = 0.95, n = 1, conf = 0.95)
```

## lrtest()
This is a function for loglikelihood ratio test of proportion. The returned value would be a list including the test statistic and p-value. An example that test the proportion of 0's to be 95\% or not:
```{r}
lrtest(c(rep(0,93),rep(1,7)),0.05)  
```

## pwls()
`pwls()` is a function to calculate the value of a special case of scoring function. It will penalize more for the underestimation than overestimation.
```{r}
pwls(0.09,0.05,0.95)
pwls(0.01,0.05,0.95)
```
Although the difference between 0.09 and 0.05 is the same as that between 0.01 and 0.05, the score for the first is smaller than the score of the second.

